---
title: "Portfolio"  # any name you want here
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(janitor)
library(raster)
library(rnaturalearth)
library(rnaturalearth)
library(tidytext)
library(textdata)
library(pdftools)
```

# Advanced Data Analysis

## Cartography in R

```{r 244 oil spills, echo=FALSE, message=FALSE, warning=FALSE}

ca_counties <- read_sf(here("data","ca_counties","CA_Counties_TIGER2016.shp"))
ca_oil_spills <- read_sf(here("data","ca_oil_spills",
                              "Oil_Spill_Incident_Tracking_%5Bds394%5D.shp")) %>% 
  clean_names()

# Filter for inland oil spills only
ca_oil_spills_inland <- ca_oil_spills %>% 
  filter(inlandmari == "Inland")

# Combining spatial layers
ca_combo <- ca_counties %>% 
  st_join(ca_oil_spills_inland)

# Count spills by county
ca_spill_counts <- ca_combo %>% count(localecoun)
  
# Create final chloropleth
ggplot(data = ca_spill_counts) +
  geom_sf(aes(fill = n), color = "black", size = 0.1) +
  scale_fill_gradientn(colors = c("light green","orange","red","black")) +
  labs(x = "Longitude", y = "Latitude", fill = "Oil spills (n)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))
```
*Chloropleth map of total inland (non-marine) oil spills by county in California in 2008 (from https://map.dfg.ca.gov/metadata/ds0394.html)*

```{r 244 cetacean raster, message=FALSE, warning=FALSE, include=FALSE}
# Create list of all cetacean rasters
cetacean_rasters <- list.files(
  here("data/ca_cetaceans"),
  full.names = TRUE
)

# Stack all rasters
cetacean_stack <- stack(cetacean_rasters)

# Write function to filter for presence threshold of 0.6
cetacean_filter <- function(x, thresh = .6){
  y <- ifelse(x >= thresh, 1, NA)
  return(y)
}

# Apply filter function to raster stack
cetacean_filtered <- calc(cetacean_stack, fun = cetacean_filter)

# Calculate total species numbers per raster cell
cetacean_sum <- calc(cetacean_filtered, fun = sum, na.rm = TRUE)

# Read in coastal outline from natural earth
pacific_coast <- ne_download(scale = 110, type = 'countries', category = 'cultural', returnclass = 'sf')

# Convert raster to df
cetacean_df <- raster::rasterToPoints(cetacean_sum) %>%
  as.data.frame()

# Create final figure
cetacean_fig <- ggplot() +
  geom_raster(data = cetacean_df, 
                       aes(x = x, y = y, fill = layer)) +
  geom_sf(data = pacific_coast, fill = "black", color = "gray") +
  coord_sf(expand= 0, xlim = c(-125, -115), ylim = c(32, 38)) +
  labs(x = "Longitude", y = "Latitude", fill = "Species richness (n)") +
  theme(legend.position = "bottom") +
  scale_fill_gradient(low = "red", high = "blue")
```

```{r fig, echo=FALSE, message=FALSE, warning=FALSE}
cetacean_fig
```
*Cetacean species richness for the Southern California Bight and surrounding areas of the Eastern Pacific Ocean. Species abundance was calculated based on a presence threshold probability of **0.6**, and rasters for all 35 species were aggregated after values over 0.6 were coerced to 1 (and values under 0.6 to 0). Highest species abundance value shown is 28 species (dark blue)*

## Text Analysis

```{r tolstoy, echo=FALSE, message=FALSE, warning=FALSE}
# Read in & tidy text
tolstoy_text <- pdf_text(here("data/war-and-peace.pdf"))

tolstoy_tidy <- data.frame(tolstoy_text) %>% 
  mutate(text_full = str_split(tolstoy_text, pattern = '\\r\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full))

# Sort data frame by chapter
tolstoy_df <- tolstoy_tidy %>% 
  slice(-(1:12)) %>% 
  mutate(chapter = case_when(
    str_detect(text_full, "Chapter") ~ text_full,
    TRUE ~ NA_character_
  )) %>% 
  fill(chapter) %>% 
  separate(col = chapter, into = c("ch", "no"), sep = " ") %>% 
  mutate(chapter = as.numeric(as.roman(no)))

# Isolate words using tokens
tolstoy_tokens <- tolstoy_df %>% 
  unnest_tokens(word, text_full) %>% 
  #select(-tolstoy_text) %>% 
  filter(is.na(as.numeric(word)))

# Remove stop words
tolstoy_nonstop_words <- tolstoy_tokens %>% 
  anti_join(stop_words)

# Word count by chapter
nonstop_counts <- tolstoy_nonstop_words %>% 
  count(chapter, word)

nonstop_counts_wp <- nonstop_counts %>% 
  filter(word == "war" | word == "peace") %>% 
  mutate(n = as.numeric(n)) %>% 
  mutate(word = as.factor(word))

war_or_peace <- nonstop_counts_wp %>% 
  group_by(chapter) %>% 
  pivot_wider(names_from = word,
               values_from = n) %>% 
  mutate(chapter = as.numeric(chapter))

ggplot() +
  geom_line(data = war_or_peace, size = 2,
            aes(x = chapter, y = war, color = "War")) + 
  geom_line(data = war_or_peace, size = 2,
            aes(x = chapter, y = peace, color = "Peace")) +
  theme_minimal() +
  ggtitle("War and Peace in War and Peace") +
  labs(x = "Chapter", y = "Word count", color = "Word") +
  scale_x_continuous(expand = c(0,0), limits = c(1, 39), breaks = seq(from = 0,
                                                                      to = 39,
                                                                      by = 3)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_manual(values=c("light blue", "red"))
```

*Prevalence of words "war" (`afinn` sentiment -2) and "peace" (`afinn` sentiment 2) by chapter in *War and Peace 


# Rhone Ecohydrology

# Forests to Faucets

Cetacean data citation: *Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org*

Analyzed text citation: *Tolstoy, Leo. War and Peace. Walter Scott Pub. Co., 1920. https://www.planetebook.com/free-ebooks/war-and-peace.pdf*